---
title: "Math678MD"
author: "Kaleigh Benesch"
date: '2023-05-31'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
```

```{r}
setwd("~/Desktop/MATH 678")

contaminant <- read.csv(file = "contaminants.csv")
# Data is at least 6 years old (April 2017). That is when it was uploaded to the site.

head(contaminant)

dim(contaminant)
# 367,289 observations and 11 predictors (will remove 'href')

names(contaminant)
```

```{r}
summary(contaminant)

contaminant <- contaminant[ ,-11] # Removes 'href'
```
Data frame with only contaminant, supplier, max, and average
```{r}
VarDescribe <- contaminant %>% select(contaminant,average_result,max_result,supplier_name)
VarDescribe
```


```{r}
# Have to figure out how to deal with not all locations having a city.

contaminant$locations_served2 <- contaminant$locations_served
contaminant <- contaminant %>% relocate(locations_served2, .after = locations_served)

contaminant1 <- contaminant %>% separate(locations_served, c('City', 'State'), sep = ", ")
contaminant1
```

```{r}
contaminant1$locations_served2 <- sub(".*? ", "", contaminant1$locations_served2)


contaminant1$locations_served2 <- sub(".* ", "", contaminant1$locations_served2)
# Removes everything before state abbreviations, so that only state is left.
contaminant1

min(nchar(contaminant1$locations_served2)) # 2 characters
max(nchar(contaminant1$locations_served2)) # 7 characters

unique(contaminant1$locations_served2) # Observations with 'CITY,' and 'LONDON,'

contaminant1 <- contaminant1[-(364733:364740), ] # 'LONDON,' observations are removed.

unique(contaminant1$locations_served2) # Observations with 'CITY,' left.
```

```{r}
CITYbruh <- filter(contaminant1, locations_served2 == "CITY,")
CITYbruh

sum(contaminant1$locations_served2 == "CITY,", na.rm = TRUE) # 9 observations with "CITY,".

contaminant1 <- contaminant1[-(129901:129909), ] # 'CITY,' observations are removed.

min(nchar(contaminant1$locations_served2)) # 2 characters
max(nchar(contaminant1$locations_served2)) # 2 characters. So, now we only have state abbreviations in the 'locations_served2' column.
```

```{r}
colnames(contaminant1)[10] = "State"
contaminant1 <- contaminant1[ ,-(9)] # Drop old state column.

min(nchar(contaminant1$City)) # 2 characters (state abbrv that didn't separate initially)
max(nchar(contaminant1$City)) # 30 characters

contaminant1
```


```{r}
# Removing the 'Yes' and 'No from columns about if limit was exceeded or not.

contaminant2 <- contaminant1 %>% mutate_at(c("health_limit_exceeded", "legal_limit_exceeded"), str_replace, "No", "") 
contaminant3 <- contaminant2 %>% mutate_at(c("health_limit_exceeded", "legal_limit_exceeded"), str_replace, "Yes", "")

colnames(contaminant3)[5] = "Health_limit"
colnames(contaminant3)[6] = "Legal_limit"

contaminant3
```

```{r}
contaminant4 <- contaminant3 %>% separate(Legal_limit, c('Legal_limit', 'Unit_measurement'), sep = " ")

contaminant4$average_result <- gsub("\\ .*", "", contaminant4$average_result)
contaminant4$max_result <- gsub("\\ .*", "", contaminant4$max_result)
contaminant4$Health_limit <- gsub("\\ .*", "", contaminant4$Health_limit)

contaminant4

# Convert the characters to numeric columns so that we can use > for ifelse() to see if an observation
# has exceeded the limits.
contaminant4$average_result <- as.numeric(contaminant4$average_result)
contaminant4$max_result <- as.numeric(contaminant4$max_result)
contaminant4$Health_limit <- as.numeric(contaminant4$Health_limit)
contaminant4$Legal_limit <- as.numeric(contaminant4$Legal_limit)

summary(contaminant4)
```

NA values for Health and Legal limits.
```{r}
sum(is.na(contaminant4$Health_limit)) # 141,052 NAs
sum(is.na(contaminant4$Legal_limit)) # 29,962 NAs
```
Fill in NA values for Legal limits.
```{r}
unique(contaminant1$legal_limit_exceeded)
# "Legal at any level. This is the Federal Limit. State Limits may be lower." causes NAs in this column now.
```

```{r}
contaminant4$Legal_limit <- replace(contaminant4$Legal_limit, is.na(contaminant4$Legal_limit), 999999)
```

Fill in NA values for Health limits.
```{r}
unique(contaminant1$health_limit_exceeded)
# "MCLG (maximum contaminant level goal): A non-enforceable health goal that is set at a level at which no known or anticipated adverse effect on the health of persons occurs and which allows an adequate margin of safety. Source: U.S. Environmental Protection Agency : 0" causes NAs in this column now.
```
MCLGs consider only public health and not the limits of detection and treatment technology effectiveness. They sometimes are set at levels which water systems cannot meet because of technological limitations.
* For microbial contaminants that may present public health risk, EPA sets the MCLG at 0. This is because ingesting one protozoan, virus, or bacterium may cause adverse health effects.
* For chemical contaminants that are carcinogens, EPA sets the MCLG at 0 if both of these are the case:
  - there is evidence that a chemical may cause cancer
  - there is no dose below which the chemical is considered safe.
* If a chemical is carcinogenic and a safe dose can be determined, EPA sets the MCLG at a level above 0 that is safe.
```{r}
contaminant4$Health_limit <- replace(contaminant4$Health_limit, is.na(contaminant4$Health_limit), 0)
```

Make new binary columns to see when limits are exceeded or not.
```{r}
contaminant4$avg_health_exceed01 <- ifelse(contaminant4$average_result > contaminant4$Health_limit, 1, 0)
contaminant4$max_health_exceed01 <- ifelse(contaminant4$max_result > contaminant4$Health_limit, 1, 0)
contaminant4$avg_legal_exceed01 <- ifelse(contaminant4$average_result > contaminant4$Legal_limit, 1, 0)
contaminant4$max_legal_exceed01 <- ifelse(contaminant4$max_result > contaminant4$Legal_limit, 1, 0)

order1 = c('id', 'contaminant', 'Unit_measurement', 'average_result', 'max_result', 'Health_limit', 'Legal_limit', 
           'avg_health_exceed01', 'max_health_exceed01', 'avg_legal_exceed01', 'max_legal_exceed01',
           'zipcode', 'City', 'State', 'supplier_name', 'number_of_people_served')
contaminant4 <- contaminant4[ , order1]
contaminant4
```

When some results are not taken.
```{r}
sum(is.na(contaminant4$avg_health_exceed01)) # 6,488 NAs
sum(is.na(contaminant4$max_health_exceed01)) # 2,938 NAs
sum(is.na(contaminant4$avg_legal_exceed01)) # 6,488 NAs
sum(is.na(contaminant4$max_legal_exceed01)) # 2,938 NAs
```

Contaminants and NA values.
```{r}
sum(is.na(contaminant4$Unit_measurement))

unique(contaminant4$contaminant[!complete.cases(contaminant4$Unit_measurement)])
# Only lead returns NA values, but the unit measurement is actually "ppb"

contaminant4$Unit_measurement <- replace(contaminant4$Unit_measurement, is.na(contaminant4$Unit_measurement), "ppb")
```


Export contaminant4
```{r}
# write.csv(contaminant4, "contaminant_updated6-7.csv", row.names = FALSE)
```


```{r}
unique(contaminant4$contaminant) # 315 different contaminants
```



EWG Drinking Water Standards Data Extraction
```{r}
standards2021 <- read.csv(file = "EWG_DWStandards.csv")
```

```{r}
standards2021 <- standards2021 %>% separate(EWG_Standard, c('EWG_standard', 'Unit_measurement'), sep = " ")

standards2021$Federal_Legal_Limit <- gsub("\\ .*", "", standards2021$Federal_Legal_Limit)

head(standards2021)
```

```{r}
unique(standards2021$Contaminant) # 99 different contaminants.

standards2021 <- standards2021[-(75), ] # Removes a duplicate contaminant and limit (p-Dichlorobenzene).
```

```{r}
colnames(standards2021)[1] = "contaminant"
colnames(standards2021)[2] = "Legal_limit2021"
colnames(standards2021)[3] = "Health_limit2021"
```

```{r}
summary(standards2021)
# Change character columns to numeric.
standards2021$Health_limit2021 <- as.numeric(standards2021$Health_limit2021)
standards2021$Legal_limit2021 <- as.numeric(standards2021$Legal_limit2021)
```

```{r}
# Turn the NA (nonexistent limits in "Legal" to large number for evaluating purposes)

standards2021$Legal_limit2021 <- replace(standards2021$Legal_limit2021, is.na(standards2021$Legal_limit2021), 999999)
```

Need to drop " (total)" from contaminant4 and standards2021 so that the "contaminant" columns join correctly.
```{r}
contaminant4$contaminant <- gsub("\\s*\\([^\\)]+\\)", "", as.character(contaminant4$contaminant))
standards2021$contaminant <- gsub("\\s*\\([^\\)]+\\)", "", as.character(standards2021$contaminant))
```

Export standards2021
```{r}
#write.csv(standards2021, "EWG2021cleaned.csv", row.names = FALSE)
```

################################################################################
Joined data set with both limits and both unit measurements.

```{r}
colnames(standards2021)[4] = "Unit_measurement2021"
```

Join standards2021 to contaminant4 according to "contaminant"
```{r}
contaminant5 <- merge(contaminant4, standards2021, by = "contaminant", all.x = TRUE)

sum(is.na(contaminant5$Legal_limit2021)) # 75,324 missing observations
sum(is.na(contaminant5$Health_limit2021)) # 75,324 missing observations

sum(is.na(contaminant5$average_result))
sum(is.na(contaminant5$max_result))
```

```{r}
order2 = c('id', 'contaminant', 'Unit_measurement', 'average_result', 'max_result', 'Health_limit', 'Health_limit2021', 
           'Legal_limit', 'Legal_limit2021', 'Unit_measurement2021', 'avg_health_exceed01', 'max_health_exceed01',
           'avg_legal_exceed01', 'max_legal_exceed01', 'zipcode', 'City', 'State', 'supplier_name', 'number_of_people_served')
contaminant5 <- contaminant5[ , order2]
```

Convert ppb to ppm
```{r}
contaminant5 <- contaminant5 %>%
  mutate(across(starts_with("Legal_limit"), 
                ~ if_else(get(str_replace(cur_column(), "Legal_limit", "Unit_measurement")) == "ppb" & !is.na(.), . / 1000, .),
                .names = "{.col}"))

contaminant5 <- contaminant5 %>%
  mutate(across(starts_with("Health_limit"), 
                ~ if_else(get(str_replace(cur_column(), "Health_limit", "Unit_measurement")) == "ppb" & !is.na(.), . / 1000, .),
                .names = "{.col}"))

contaminant5 <- contaminant5 %>%
  mutate(across(ends_with("average_result"), 
                ~ if_else(get(str_replace(cur_column(), "average_result", "Unit_measurement")) == "ppb" & !is.na(.), . / 1000, .),
                .names = "{.col}"))

contaminant5 <- contaminant5 %>%
  mutate(across(ends_with("max_result"), 
                ~ if_else(get(str_replace(cur_column(), "max_result", "Unit_measurement")) == "ppb" & !is.na(.), . / 1000, .),
                .names = "{.col}"))

contaminant5$Unit_measurement[contaminant5$Unit_measurement == "ppb"] <- "ppm"
contaminant5$Unit_measurement2021[contaminant5$Unit_measurement2021 == "ppb"] <- "ppm"

contaminant5
```

Unit measurements with "at" instead of a real unit.
```{r}
sum(contaminant5$Unit_measurement == "at", na.rm = TRUE) # 14,270 observations with "at" for unit measurement.

at_units <- filter(contaminant5, Unit_measurement == "at")
at_units

# Not really important, so we will just turn them to NAs...they just got lost among data cleaning.
contaminant5$Unit_measurement[contaminant5$Unit_measurement == "at"] = NA
``` 

Final cleaning, make accidental state abbreviations in the "City" column NAs.
```{r}
contaminant5$City <- gsub(nchar(contaminant5$State) == 2, NA, contaminant5$City)
contaminant5$City <- replace(contaminant5$City, nchar(contaminant5$City) == 2, NA)

min(nchar(contaminant5$City))
sum(is.na(contaminant5$City)) # 92,337 missing "City" values.
``` 


Export contaminant5 (joined data frame)
```{r}
# write.csv(contaminant5, "contaminant_joined6-8.csv", row.names = FALSE)
```


################################################################################

Answering the first proposal question: Are some suppliers of drinking water more likely to exceed legal limits of contaminants than other suppliers?


Suppliers and contaminants.
```{r}
length(unique(contaminant5$supplier_name)) # 42,668 unique suppliers
```

```{r}
supplier <- contaminant5 %>% select('avg_health_exceed01', 'avg_legal_exceed01', 
                                    'max_health_exceed01', 'max_legal_exceed01', 'supplier_name', 'State', 'contaminant')

# How many times total a limit is not and is exceeded.
table(supplier$avg_legal_exceed01)
table(supplier$max_legal_exceed01)
table(supplier$avg_health_exceed01)
table(supplier$max_health_exceed01)
```

```{r}
supplier_freq <- table(supplier$supplier_name) # Times each supplier occurs.
head(supplier_freq)

# Turn table into a data frame
supplier_freq = as.data.frame(supplier_freq)
head(supplier_freq)

supplier2 <- aggregate(cbind(avg_health_exceed01, max_health_exceed01, avg_legal_exceed01, max_legal_exceed01) ~ supplier_name, supplier, sum, na.rm = FALSE) 
# Times each supplier average/max sample exceeds health/legal limit.
```

```{r}
length(unique(supplier2$supplier_name)) # has 42,566 suppliers
length(unique(supplier_freq$Var1)) # has all suppliers : 42,668

missing <- supplier_freq %>% filter(!supplier_freq$Var1 %in% supplier2$supplier_name)
head(missing) # 102 missing suppliers

supplier_freq <- anti_join(supplier_freq, missing)

length(unique(supplier_freq$Var1)) # now they are both the same length

colnames(supplier_freq)[1] = "supplier_name"
```

```{r}
supplier3 <- merge(supplier_freq, supplier2, by = "supplier_name", all.x = TRUE)
head(supplier3)

colnames(supplier3)[2] = "total_samples"
colnames(supplier3)[3] = "avg_health_exceed"
colnames(supplier3)[4] = "max_health_exceed"
colnames(supplier3)[5] = "avg_legal_exceed"
colnames(supplier3)[6] = "max_legal_exceed"
```

Make columns for rate of times exceeded.
```{r}
supplier3$rate_avg_H_exceed <- (supplier3$avg_health_exceed / supplier3$total_samples)
supplier3$rate_max_H_exceed <- (supplier3$max_health_exceed / supplier3$total_samples)
supplier3$rate_avg_L_exceed <- (supplier3$avg_legal_exceed / supplier3$total_samples)
supplier3$rate_max_L_exceed <- (supplier3$max_legal_exceed / supplier3$total_samples)
```

** Maybe group certain ratings and create a bar graph of count. **
```{r}
supplier4 <- supplier3 %>% 
  mutate(max_health_grouped = ifelse(rate_max_H_exceed <= 0.25, "Low",
                                     ifelse(rate_max_H_exceed >= 0.25 & rate_max_H_exceed <= 0.50, "Medium",
                                            ifelse(rate_max_H_exceed >= 0.50 & rate_max_H_exceed <= 0.75, "High",
                                                   ifelse(rate_max_H_exceed >= 0.75, "Very High", NA)))))

supplier4 <- supplier4 %>% 
  mutate(max_legal_grouped = ifelse(rate_max_L_exceed <= 0.25, "Low",
                                     ifelse(rate_max_L_exceed >= 0.25 & rate_max_L_exceed <= 0.50, "Medium",
                                            ifelse(rate_max_L_exceed >= 0.50 & rate_max_L_exceed <= 0.75, "High",
                                                   ifelse(rate_max_L_exceed >= 0.75, "Very High", NA)))))

supplier4
```

Suppliers who exceed legal limits 100% of the time.
```{r}
AlwaysExceed <- filter(supplier4, rate_avg_L_exceed == 1)
AlwaysExceed

unique(AlwaysExceed$total_samples)
unique(AlwaysExceed$supplier_name)
```


```{r}
bar_order = c("Low", "Medium", "High","Very High")

plot1 <- ggplot(data = supplier4) + geom_bar(aes(x = factor(max_health_grouped, bar_order)), 
                                    color = 'lightblue', fill = 'lightblue') + 
  labs(title = "Suppliers' Maximum Sample Exceeds Health Limit", 
       x = "Rate That Suppliers Exceed Limit", 
       y = "Count")

plot2 <- ggplot(data = supplier4) + geom_bar(aes(x = factor(max_legal_grouped, bar_order)), 
                                    color = 'lightgreen', fill = 'lightgreen') + 
  labs(title = "Suppliers' Maximum Sample Exceeds Legal Limit", 
       x = "Rate That Suppliers Exceed Limit", 
       y = "Count")

plot(plot1)
plot(plot2)
```

Try as histograms instead.
```{r}
plot1.2 <- ggplot(data = supplier4) +
  geom_histogram(mapping = aes(x = rate_max_H_exceed), binwidth = 0.1, color = 'lightblue', fill = 'lightblue') + 
  labs(title = "Suppliers' Maximum Sample Exceeds Health Limit", 
       x = "Rate That Suppliers Exceed Limit", 
       y = "Count")

plot2.2 <- ggplot(data = supplier4) +
  geom_histogram(mapping = aes(x = rate_max_L_exceed), binwidth = 0.1, color = 'lightgreen', fill = 'lightgreen') + 
  labs(title = "Suppliers' Maximum Sample Exceeds Legal Limit", 
       x = "Rate That Suppliers Exceed Limit", 
       y = "Count")

plot(plot1.2)
plot(plot2.2)
```


################################################################################

Answering the second proposal question: Are there any correlated contaminants? In other words, if a specific 
contaminant is present at a certain level, is it likely that another specific contaminant is present?

The most commonly occurring contaminant samples are:
```{r}
summary(as.factor(contaminant5$contaminant))
names(sort(summary(as.factor(contaminant5$contaminant)), decreasing = TRUE)[1:8])
# "Nitrate", "Total trihalomethanes", "Barium", "Copper", "Chloroform", "Total haloacetic acids", "Bromodichloromethane",
# "Dibromochloromethane"
```

Make a more manageable data frame.
```{r}
contaminant5.2 <- contaminant5 %>% filter(contaminant == c("Nitrate", "Total trihalomethanes", 
                                                           "Barium", "Copper", "Chloroform", "Total haloacetic acids",
                                                           "Bromodichloromethane", "Dibromochloromethane"))
contaminant5.2 <- contaminant5.2 %>% select(contaminant, State, average_result)
contaminant5.2
```
Go by supplier instead of State.
```{r}
contaminant5.21 <- contaminant5 %>% filter(contaminant == c("Nitrate", "Total trihalomethanes", 
                                                           "Barium", "Copper", "Chloroform", "Total haloacetic acids",
                                                           "Bromodichloromethane", "Dibromochloromethane"))
contaminant5.21 <- contaminant5.21 %>% select(contaminant, supplier_name, average_result)
contaminant5.21
```

Average result from each State by top 8 most common contaminants.
```{r}
contaminant5.2 <- contaminant5.2 %>% 
  pivot_wider(names_from = contaminant, values_from = average_result, values_fn = mean)
contaminant5.2
```
Average result from each supplier by top 8 most common contaminants.
```{r}
contaminant5.21 <- contaminant5.21 %>% 
  pivot_wider(names_from = contaminant, values_from = average_result, values_fn = mean)
contaminant5.21
```

Correlation between contaminants with states?
```{r}
library(psych)

corPlot((contaminant5.2[2:9]), cex = 1, cex.axis = 0.8, xlas = 2, main = "Correlation Plot of 8 Most Common Contaminants")
```
While there is not strong correlation between any of these top 8 most common contaminants, the 
most correlated contaminants are "Total haloacetic acids" with "Total trihalomethanes". These two
have a positive correlation coefficient of 0.67 and might be more likely to be found together in similar
amounts when testing water supplies.

Correlation between contaminants with suppliers?
```{r}
corPlot((contaminant5.21[2:9]), cex = 1, cex.axis = 0.8, xlas = 2, main = "Correlation Plot of 8 Most Common Contaminants")
```

```{r}
summary(contaminant5.21)
sum(is.na(contaminant5.21))
```


Use a model that predicts amounts of a certain contaminant based on the others.
```{r}
# set.seed(22)
# 
# train = sample(1:nrow(contaminant5.21), 0.7*nrow(contaminant5.21))
# contaminant5.21.train = contaminant5.21[train, ]
# contaminant5.21.test = contaminant5.21[-train, ]
```

Make a liner model
```{r}
# lm.fit1 = lm(Barium ~ . - supplier_name, data = contaminant5.21.train)
# summary(lm.fit1)
```


################################################################################

Answering the third proposal question: Are some water contaminants more common in western states vs eastern states? 
Northern vs southern?

* Should I make the scales such that the top limit is the legal limit?

Install package for mapping.
```{r}
library(usmap)
```

Since there are over 300 contaminants, we will look at the top 5 most common contaminants in the data set.
We will create a map of the U.S. that shows the average amounts of each contaminant by state. The states
with higher average amounts of each contaminant will have a darker shade while lesser amounts will be
represented by light shades or white. States without a recorded average will be filled in gray. There are no
observations from Kansas, Mississippi, or Georgia.
```{r}
names(sort(summary(as.factor(contaminant5$contaminant)), decreasing = TRUE)[1:5])
# Nitrate, Total trihalomethanes, Barium, Copper, and Chloroform are the 5 most frequently appearing contaminants.
```

Nitrate
```{r}
Nitrate <- contaminant5 %>% filter(contaminant == "Nitrate")
Nitrate <- data.frame(state = Nitrate$State, 
                             average_result = Nitrate$average_result, contaminant = Nitrate$contaminant)

NitrateAvg <- aggregate(average_result ~ state, Nitrate, mean)

max(NitrateAvg$average_result) # 5.58

plot_usmap(data = NitrateAvg, values = "average_result", labels = TRUE) + 
  scale_fill_continuous(low = "white", high = "darkred", name = "average_result", 
                        label = scales::comma, limits = c(0, 6)) + 
  labs(title = "United States", subtitle = "Average Results of Nitrate (ppm) by State") +
  theme(legend.position = "right")
```

Total trihalomethanes
```{r}
Total_trihalomethanes <- contaminant5 %>% filter(contaminant == "Total trihalomethanes")
Total_trihalomethanes <- data.frame(state = Total_trihalomethanes$State, 
                             average_result = Total_trihalomethanes$average_result, 
                             contaminant = Total_trihalomethanes$contaminant)

Total_trihalomethanesAvg <- aggregate(average_result ~ state, Total_trihalomethanes, mean)

max(Total_trihalomethanesAvg$average_result) # 0.055

plot_usmap(data = Total_trihalomethanesAvg, values = "average_result", labels = TRUE) + 
  scale_fill_continuous(low = "white", high = "darkgreen", name = "average_result", 
                        label = scales::comma, limits = c(0, 0.06)) + 
  labs(title = "United States", subtitle = "Average Results of Total Trihalomethanes (ppm) by State") +
  theme(legend.position = "right")
```

Barium
```{r}
Barium <- contaminant5 %>% filter(contaminant == "Barium")
Barium <- data.frame(state = Barium$State, 
                             average_result = Barium$average_result, 
                             contaminant = Barium$contaminant)

BariumAvg <- aggregate(average_result ~ state, Barium, mean)

max(BariumAvg$average_result) # 0.298

plot_usmap(data = BariumAvg, values = "average_result", labels = TRUE) + 
  scale_fill_continuous(low = "white", high = "darkblue", name = "average_result", 
                        label = scales::comma, limits = c(0, 0.3)) + 
  labs(title = "United States", subtitle = "Average Results of Barium (ppm) by State") +
  theme(legend.position = "right")
```

Copper
```{r}
Copper <- contaminant5 %>% filter(contaminant == "Copper")
Copper <- data.frame(state = Copper$State, 
                             average_result = Copper$average_result, 
                             contaminant = Copper$contaminant)

CopperAvg <- aggregate(average_result ~ state, Copper, mean)

max(CopperAvg$average_result) # 0.353

plot_usmap(data = CopperAvg, values = "average_result", labels = TRUE) + 
  scale_fill_continuous(low = "white", high = "darkorange", name = "average_result", 
                        label = scales::comma, limits = c(0, 0.37)) + 
  labs(title = "United States", subtitle = "Average Results of Copper (ppm) by State") +
  theme(legend.position = "right")
```

Chloroform
```{r}
Chloroform <- contaminant5 %>% filter(contaminant == "Chloroform")
Chloroform <- data.frame(state = Chloroform$State, 
                             average_result = Chloroform$average_result, 
                             contaminant = Chloroform$contaminant)

ChloroformAvg <- aggregate(average_result ~ state, Chloroform, mean)

max(ChloroformAvg$average_result) # 0.029

plot_usmap(data = ChloroformAvg, values = "average_result", labels = TRUE) + 
  scale_fill_continuous(low = "white", high = "gold", name = "average_result", 
                        label = scales::comma, limits = c(0, 0.03)) + 
  labs(title = "United States", subtitle = "Average Results of Chloroform (ppm) by State") +
  theme(legend.position = "right")
```

After mapping the level of these 5 contaminants, the only one that seems to show any pattern is
Copper. According to the map results in orange, Copper mostly appears in higher amounts in 
northeastern parts of the United States and in smaller amounts in west/southwestern states.


Export contaminant5 (joined data frame)
First deadline.
```{r}
# write.csv(contaminant5, "contaminant_6-21.csv", row.names = FALSE)
```

------------------------ FIRST DEADLINE 6/21/2023 ------------------------ 






 
